\section{Convolutional neural networks}

Computer vision applied to images (stationary or dynamic), with the goal of
classifying images, detectinb objects and their locations, making predictions,
etc.

Traditional approaches use feature extraction and decision. Issues:
\begin{itemize}
		\item Finding suitable features (suitable for a machine!) is hard
		\item Highly variable data
\end{itemize}

Using fully connected neural networks:
\begin{itemize}
		\item Input is 2D image with spatial information
		\item But fully connected network has every neuron in hidden network
				connected to every neuron in input. Spatial information is
				lost.
		\item And a lot of parameters
\end{itemize}

Approach: Feature extraction with convolution.

\subsection{CNNs for classification}

\begin{enumerate}
		\item Apply convolution filters to generate feature maps
		\item Introduce non-linearity, often with ReLU
		\item Pooling: Downsampling on each feature map
\end{enumerate}

Specifically, each neuron gets only a $x \times y$ window as input.

\subsubsection{Pooling}

Max pool $\max()$ over each pool allows focusing on `positive' aspects of
features.

\subsubsection{Chaining}

Can be chained.

\begin{itemize}
		\item Repeatedly extract features via convolution, activate via ReLU, pooling
		\item Then, flatten feature vectors and feed into fully-connected layer
				for classification
		\item Output is probability of image belonging to a particular class.
\end{itemize}

\subsubsection{R-CNN}

Problem of naive usage of CNNs for object detection: CNN is going to find way
too many objects in each image. Solution: Object detection with R-CNN.

\begin{itemize}
		\item R-CNN algorithm finds regions which might have objects
		\item CNN used on those regions (warped to required dimensions) to find
				and classify objects.
\end{itemize}

\subsection{Semantic segmentation: Fully convolutional network}

\begin{itemize}
		\item FCN is CNN designed with all convolutional layers, including
				downsampling (beginning) and upsampling (end) operations
		\item Model is E2E-trained without human labelling
		\item Pre-trained models can be distributed, requiring only fine-tuning
				with application-specific data.
\end{itemize}
